---
title: "Preterm Exploratory"
author: "Peter Stella"
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(ROCR)
library(pROC)
library(MASS)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(ggfortify)
library(glmnet)
library(e1071)
library(mlbench)
library(tidyverse)
library(ISLR)
library(tree)
library(randomForest)
library(gbm)
library(ranger)
library(Hmisc)
```


```{r}

#Load data and modify column names
load('C:\\Users\\pstel\\OneDrive\\Desktop\\MachineLearning2019\\Preterm Birth\\HTA20_RMA.RData')
ano=read.csv("C:\\Users\\pstel\\OneDrive\\Desktop\\MachineLearning2019\\Preterm Birth\\anoSC1_v11_nokey.csv",stringsAsFactors = FALSE)
data_df <- as.data.frame(eset_HTA20)
annotations <- ano #%>% column_to_rownames(., var= "SampleID")
data_transposed <- as.data.frame(t(data_df))
data_transposed <- rownames_to_column(data_transposed)
data_transposed <- dplyr::rename(data_transposed, SampleID = rowname)
data_annotated <- inner_join(annotations,data_transposed)
data_2rownames <- data_annotated %>% column_to_rownames(.,var = "SampleID")
data_train <- filter(data_annotated, Train == 1)  
data_test <- filter(data_annotated, Train == 0)
data_train_rownames <- filter(data_2rownames, Train == 1)  
data_test_rownames <- filter(data_2rownames, Train == 0)  

# select only data from first batch
#data_train1 <- data_train[1:335,]
data_train1 <- data_train
data_train1 <- data_train1 %>% dplyr::select(-"Batch",-"Set",-"Train", -"Platform")
data_train1 <- data_train1 %>% mutate(term = ifelse(GA<37,0,1))
data_train1 <- data_train1 %>% dplyr::select("SampleID", "GA", "term", everything())
data_train_term <- data_train1 %>% dplyr::select(-"GA")
data_train_GA <- data_train1 %>% dplyr::select(-"term")

#data_train_rownames1 <- data_train_rownames[1:335,]
data_train_rownames1 <- data_train_rownames
data_train_rownames1 <- data_train_rownames1 %>% dplyr::select(-"Batch",-"Set",-"Train", -"Platform")
data_train_rownames1 <- data_train_rownames1 %>% mutate(term = ifelse(GA<37,0,1))
data_train_rownames1 <- data_train_rownames1 %>% dplyr::select( "GA", "term", everything())
data_train_rownames_term <- data_train_rownames1 %>% dplyr::select(-"GA")
data_train_rownames_GA <- data_train_rownames1 %>% dplyr::select(-"term")

#ctrl_test <- trainControl(method = "boot", 25)
#test_lasso <- train(GA~ ., data =data_train_GA, method = 'lasso', trControl= ctrl_test)
```




```{r}
data_test_rownames1 <- data_test_rownames %>% dplyr::select(-"Batch",-"Set",-"Train", -"Platform")
data_test_rownames1 <- data_test_rownames1 %>% mutate(term = ifelse(GA<37,0,1))
data_test_rownames1 <- data_test_rownames1 %>% dplyr::select( "GA", "term", everything())
data_test_rownames_term <- data_test_rownames1 %>% dplyr::select(-"GA")
data_test_rownames_GA <- data_test_rownames1 %>% dplyr::select(-"term")
```


```{r}
data_test_rownames_GA_label <- data_test_rownames_GA %>% select("GA")
data_test_rownames_term_label <-data_test_rownames_term %>% select("term")
data_test_rownames_GA_no_label <- data_test_rownames_GA %>% select(-"GA")
data_test_rownames_term_no_label <- data_test_rownames_term %>% select(-"term")
GA_test_no_lab_mat <- data.matrix(data_test_rownames_GA_no_label, rownames.force = TRUE)
GA_test_lab_mat <- data.matrix(data_test_rownames_GA_label, rownames.force = TRUE)
```



```{r}
#convert numeric scoring to factor for term
data_train_term <- data_train_term %>% mutate(term = ifelse(term==1, "yes", "no"))
data_train_term$term <- as.factor(data_train_term$term)

```


Now we perform feature selection with Fselector, FESER, and PCA
#PCA

```{r}
data_pca_GA <- data_train1 %>% dplyr::select(-"SampleID", -"GA", -"term")
pca_GA <- prcomp(data_pca_GA, scale. = TRUE)
```

```{r}
fviz_eig(pca_GA, ncp=25)
```

#FSelector

```{r}
lct <- linear.correlation(GA~., data_train_rownames_GA)
chi <-chi.squared(term~., data_train_rownames_term)
```



FESER
#needs matrices with separate objects so 

```{r}
data_train_rownames_GA_label <- data_train_rownames_GA %>% select("GA")
data_train_rownames_term_label <-data_train_rownames_term %>% select("term")
data_train_rownames_GA_no_label <- data_train_rownames_GA %>% select(-"GA")
data_train_rownames_term_no_label <- data_train_rownames_term %>% select(-"term")
GA_no_lab_mat <- data.matrix(data_train_rownames_GA_no_label, rownames.force = TRUE)
GA_lab_mat <- data.matrix(data_train_rownames_GA_label, rownames.force = TRUE)
GA_no_lab_mat2 <- GA_no_lab_mat[,1:32800]

```


```{r}
ridge_fit_cv_nofilt = cv.glmnet(x = scale(GA_no_lab_mat2), y = GA_lab_mat, alpha = 0)

```

```{r}
ridge_fit_nofilt  <-  glmnet(x = scale(GA_no_lab_mat2), y = GA_lab_mat, alpha = 0)
```


```{r}
ridge_fit_cv_nofilt
```


```{r}
sqrt(ridge_fit_cv_nofilt$cvm[which.min(abs(ridge_fit_cv_nofilt$lambda - ridge_fit_cv_nofilt$lambda.1se))])

```


```{r}
feser_cor <- as.data.frame(filter.corr(features= GA_no_lab_mat, class = GA_lab_mat, mincorr = 0.1))


```


```{r}
inf_gain_filt <- filter.gain.inf(features = GA_no_lab_mat, class = GA_lab_mat, zero.gain.out = TRUE)
```


Feature selection performed: 
Now need build datasets. Broadly, we take our features and combine with the original data
```{r}
data_train_GA_fescor <- data.frame(data_train_GA$GA,feser_cor)
data_train_GA_fescor <- dplyr::rename(data_train_GA_fescor, GA= data_train_GA.GA )
```





```{r}
#inf_features <- colnames(inf_gain_filt)
#data_train_GA_inf_top10 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= inf_features))
#data_train_GA_inf_top10 <- dplyr::rename(data_train_GA_inf_top10, GA = "dat in_term", select= inf_features)

```



#From F selector


```{r}
ggplot(lctest_ordered_row,aes(x=rowname, y= attr_importance))+geom_point()+xlim(0,30000)
  
  
#scale_x_continuous(limits=c(0, 10))
```




```{r}
lct <- rownames_to_column(lct)
lct <- dplyr::rename(lct, gene = rowname)
lctest_ordered <- lct %>% arrange(desc(attr_importance))
lct_top100 <- lctest_ordered[1:100,]
lct_top200 <- lctest_ordered[1:200,]
lct_top500 <- lctest_ordered[1:500,]
lct_top1000 <- lctest_ordered[1:1000,]
```

ordered lists of genes, subsets of database by tops 

```{r}
lct_100 <- lct_top100$gene
lct_200 <- lct_top200$gene
lct_500 <- lct_top500$gene
lct_1000 <- lct_top1000$gene

data_train_GA_lc_top100 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= lct_100))
data_train_GA_lc_top100 <- dplyr::rename(data_train_GA_lc_top100, GA= data_train_GA.GA )
data_train_GA_lc_top500 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= lct_500))
data_train_GA_lc_top500 <- dplyr::rename(data_train_GA_lc_top500, GA= data_train_GA.GA )
data_train_GA_lc_top1000 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= lct_1000))
data_train_GA_lc_top1000 <- dplyr::rename(data_train_GA_lc_top1000, GA= data_train_GA.GA )


```


```{r}
#chi <- rownames_to_column(chi)
##chi <- dplyr::rename(chi, gene = rowname)
#chitest_ordered <- chi %>% arrange(desc(attr_importance))
#chi_top100 <- chitest_ordered[1:100,]
#chi_top200 <- chitest_ordered[1:200,]
#chi_top500 <- chitest_ordered[1:500,]
#chi_top1000 <- chitest_ordered[1:1000,]
```


```{r}
#chi_100 <- chi_top100$gene
#chi_200 <- chi_top200$gene
#chi_500 <- chi_top500$gene
#chi_1000 <- chi_top1000$gene

#data_train_term_chi_top100 <- data.frame(data_train_term$term, subset(data_train_term, select= chi_100))
#data_train_term_chi_top100 <- dplyr::rename(data_train_term_chi_top100, term= data_train_term.term )
#data_train_term_chi_top500 <- data.frame(data_train_term$term, subset(data_train_term, select= chi_500))
#data_train_term_chi_top500 <- dplyr::rename(data_train_term_chi_top500, term= data_train_term.term )
#data_train_term_chi_top1000 <- data.frame(data_train_term$term, subset(data_train_term, select= chi_1000))
#data_train_term_chi_top1000 <- dplyr::rename(data_train_term_chi_top1000, term= data_train_term.term )

```

From PCA

```{r}
pca_5 <- pca_GA$x[,1:5]
pca_25 <- pca_GA$x[,1:25]
pca_100 <- pca_GA$x[,1:100]
```

```{r}
data_train_GA_pca_5 <- as.data.frame(cbind(data_train_GA$GA, pca_5))
data_train_GA_pca_5 <- dplyr::rename(data_train_GA_pca_5, GA = V1)

data_train_GA_pca_25 <- as.data.frame(cbind(data_train_GA$GA, pca_25))
data_train_GA_pca_25 <- dplyr::rename(data_train_GA_pca_25, GA = V1)

data_train_GA_pca_100 <- as.data.frame(cbind(data_train_GA$GA, pca_100))
data_train_GA_pca_100 <- dplyr::rename(data_train_GA_pca_100, GA = V1)

```

```{r}
pca_GA
```


#PCA

1. Regression against age

# Caret control funtions: 

```{r}
reg_control <-  trainControl(method = "repeatedcv", repeats = 5, number = 5)
class_control <- trainControl(method= "repeatedcv", repeats = 5, number = 5, classProbs = T, savePredictions = T)
```

#Correlation filter from FSelector
```{r}
FS_reg_100 <- caret::train(GA~., data=data_train_GA_lc_top100, method = "glm", family = "gaussian", trControl= reg_control)
FS_reg_500 <- caret::train(GA~., data=data_train_GA_lc_top500, method = "glm", family = "gaussian", trControl= reg_control)
```
WRONG

#Feser correlation
```{r}
Feser_reg_cor <- caret::train(GA~., data=data_train_GA_fescor, method = "glm", family = "gaussian", trControl= reg_control)
```



#Chi squared
```{r}
term_chi100_class <- caret::train(term~., data=data_train_term_chi_top100, method= "glm", family= "binomial", trControl = class_control)
term_chi500_class <- caret::train(term~., data=data_train_term_chi_top500, method= "glm", family= "binomial", trControl = class_control)

```
#information loss
```{r}
#inf_reg <- caret::train(GA~., data=data_train_GA_inf_top10, method = "glm", family = "gaussian", trControl= reg_control)
#inf_class <- caret::train(term~., data=data_train_term_inf_top10, method = "glm", family = "binomial", trControl= class_control)
```
#PCA
```{r}
pca_reg_5 <- caret::train(GA~., data=data_train_GA_pca_5, method = "glm", family = "gaussian", trControl= reg_control)
pca_reg_25<- caret::train(GA~., data=data_train_GA_pca_25, method = "glm", family = "gaussian", trControl= reg_control)
pca_reg_100<- caret::train(GA~., data=data_train_GA_pca_100, method = "glm", family = "gaussian", trControl= reg_control)
```

```{r}

regression_models <- c("FS_100", "FS_500", "Feser.1", "Feser.inf","PCA_5", "PCA_25", "PCA_100")
regression_RMSE <- c(FS_reg_100$results$RMSE, FS_reg_500$results$RMSE, Feser_reg_cor$results$RMSE,inf_reg$results$RMSE,pca_reg_5$results$RMSE,pca_reg_25$results$RMSE, pca_reg_100$results$RMSE)
regression_sum <- cbind(regression_models, regression_RMSE)
```


```{r}
RF_reg100 <- caret::train(GA~., data=data_train_GA_rf_top100, method = "glm", family = "gaussian", trControl= reg_control)
RF_reg500 <- caret::train(GA~., data=data_train_GA_rf_top500, method = "glm", family = "gaussian", trControl= reg_control)
RF_reg1000 <- caret::train(GA~., data=data_train_GA_rf_top1000, method = "glm", family = "gaussian", trControl= reg_control)

```



```{r}
RF_reg100
RF_reg500
RF_reg1000
```

```{r}
regression_sum
```




Not so hot, lets try more sophisticated regression techniques, like SVM


```{r}
str(data_train_GA_lc_top100)
```

```{r}
reg_control <-  trainControl(method = "repeatedcv", repeats = 5, number = 5)
class_control <- trainControl(method= "repeatedcv", repeats = 5, number = 5, classProbs = T, savePredictions = T)
```

```{r}
SVM_GA_100 <- caret::train(GA~., data=data_train_GA_lc_top100, method= "svmLinear", trControl = reg_control)
SVM_GA_500 <- caret::train(GA~., data=data_train_GA_lc_top500, method= "svmLinear", trControl = reg_control)

```

```{r}
SVM_rf_100 <- caret::train(GA~., data=data_train_GA_rf_top100, method= "svmLinear", trControl = reg_control)

SVM_rf_500 <- caret::train(GA~., data=data_train_GA_rf_top500, method= "svmLinear", trControl = reg_control)

SVM_rf_1000 <- caret::train(GA~., data=data_train_GA_rf_top1000, method= "svmLinear", trControl = reg_control)
```


```{r}
SVM_rf_r_100 <- caret::train(GA~., data=data_train_GA_rf_top100, method= "svmRadial", trControl = reg_control)

SVM_rf_r_500 <- caret::train(GA~., data=data_train_GA_rf_top500, method= "svmRadial", trControl = reg_control)

SVM_rf_r_1000 <- caret::train(GA~., data=data_train_GA_rf_top1000, method= "svmRadial", trControl = reg_control)
```



```{r}
SVM_rf_100
SVM_rf_r_100
SVM_rf_500
SVM_rf_r_500
SVM_rf_1000
SVM_rf_r_1000

```




```{r}
SVM_GA_Feser <- caret::train(GA~., data=data_train_GA_fescor, method= "svmLinear", trControl = reg_control)

```

```{r}
SVM_GA_inf<- caret::train(GA~., data=data_train_GA_inf_top10, method= "svmLinear", trControl = reg_control)
```


```{r}
SVM_GA_pca_5 <- caret::train(GA~., data=data_train_GA_pca_5, method= "svmLinear", trControl = reg_control)
SVM_GA_pca_100 <- caret::train(GA~., data=data_train_GA_pca_100, method= "svmLinear", trControl = reg_control)
```

*FS_reg_100
FS_reg_500
*Feser_reg_10
Feser_reg_100
*Feser_reg_1000
*inf_reg
*pca_reg_5
pca_reg_25
*pca_reg_100
```{r}
SVM_GA_100
SVM_GA_500
SVM_GA_Feser
SVM_GA_inf
SVM_GA_pca_5
SVM_GA_pca_100
```


Now try radial 


```{r}
reg_control <-  trainControl(method = "repeatedcv", repeats = 5, number = 5)
class_control <- trainControl(method= "repeatedcv", repeats = 5, number = 5, classProbs = T, savePredictions = T)
```

```{r}
SVM_r_GA_100 <- caret::train(GA~., data=data_train_GA_lc_top100, method= "svmRadial", trControl = reg_control)
SVM_r_GA_500 <- caret::train(GA~., data=data_train_GA_lc_top500, method= "svmRadial", trControl = reg_control)
SVM_r_GA_1000 <- caret::train(GA~., data=data_train_GA_lc_top1000, method= "svmRadial", trControl = reg_control)
```

```{r}
SVM_r_GA_Feser <- caret::train(GA~., data=data_train_GA_fescor, method= "svmRadial", trControl = reg_control)

```

```{r}
SVM_r_GA_inf<- caret::train(GA~., data=data_train_GA_inf_top10, method= "svmRadial", trControl = reg_control)
```

```{r}
SVM_r_GA_pca_5 <- caret::train(GA~., data=data_train_GA_pca_5, method= "svmRadial", trControl = reg_control)
SVM_r_GA_pca_100 <- caret::train(GA~., data=data_train_GA_pca_100, method= "svmRadial", trControl = reg_control)
```


```{r}
SVM_r_GA_100
SVM_GA_100
SVM_r_GA_500
SVM_GA_500
SVM_r_GA_Feser
SVM_GA_Feser
SVM_r_GA_pca_5
SVM_r_GA_pca_100
SVM_r_GA_1000
```

Still not so hot. How about random forest?

```{r}
control_rf_term <- trainControl(method="repeatedcv", number = 5, repeats = 5, classProbs = T, savePredictions = T)
control_rf_GA <- trainControl(method="repeatedcv", number = 5, repeats = 5)

```


```{r}
rf_term_100 <- caret::train(term~., data= data_train_term_chi_top100, method = "rf", trControl = control_rf_term)
rf_term_500 <- caret::train(term~., data= data_train_term_chi_top500, method = "rf", trControl = control_rf_term)

```

```{r}
rf_GA_100 <- caret::train(GA~., data= data_train_GA_lc_top100, method = "rf", trControl = control_rf_GA)
rf_GA_500 <- caret::train(GA~., data= data_train_GA_lc_top500, method = "rf", trControl = control_rf_GA)
rf_GA_1000 <- caret::train(GA~., data= data_train_GA_lc_top1000, method = "rf", trControl = control_rf_GA)

```

```{r}
rf_feser <- caret::train(GA~.RF, data= data_train_GA_fescor, method="ranger", trControl=control_rf_GA)
```


```{r}
rf_term_inf <- caret::train(term~., data= data_train_term_inf_top10, method = "rf", trControl = control_rf_term)
rf_GA_inf <- caret::train(GA~., data= data_train_GA_inf_top10, method = "rf", trControl = control_rf_GA)
```


```{r}
rf_GA_pca_5 <- caret::train(GA~., data= data_train_GA_pca_5, method = "rf", trControl = control_rf_GA)
rf_GA_pca_25 <- caret::train(GA~., data= data_train_GA_pca_25, method = "rf", trControl = control_rf_GA)
rf_GA_pca_100 <-  caret::train(GA~., data= data_train_GA_pca_100, method = "rf", trControl = control_rf_GA) 
```


```{r}
ranger_kaboodle <- ranger(dependent.variable.name="GA", data= data_train_rownames_GA, importance = "impurity",  num.trees= 10000, mtry = 500)

```



```{r}
rf_GA_100
rf_GA_500
rf_GA_1000
rf_GA_pca_5
rf_GA_pca_100
rf_feser
ranger_kaboodle
```


```{r}
sqrt(53.76)
```


```{r}
ranger_kaboodle
rang_imps <- ranger::importance(ranger_kaboodle)
```


```{r}
df_rimps <- as.data.frame(rang_imps)
```


```{r}
#df_rimps <- rownames_to_column(df_rimps)
#df_rimps <- dplyr::rename(df_rimps, Gene = rowname )
df_rimp_ordered<- df_rimps %>% arrange(desc(rang_imps))
str(df_rimp_ordered)
```




```{r}
rf_top100 <- df_rimp_ordered[1:100,]
rf_top100 <- rf_top100$Gene
rf_top500 <- df_rimp_ordered[1:500,]
rf_top500 <- rf_top500$Gene
rf_top1000 <- df_rimp_ordered[1:1000,]
rf_top1000 <- rf_top1000$Gene

```


```{r}
data_train_GA_rf_top100 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= rf_top100))
data_train_GA_rf_top100 <- dplyr::rename(data_train_GA_rf_top100, GA= data_train_GA.GA )
data_train_GA_rf_top500 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= rf_top500))
data_train_GA_rf_top500 <- dplyr::rename(data_train_GA_rf_top500, GA= data_train_GA.GA)
data_train_GA_rf_top1000 <- data.frame(data_train_GA$GA, subset(data_train_GA, select= rf_top1000))
data_train_GA_rf_top1000 <- dplyr::rename(data_train_GA_rf_top1000, GA= data_train_GA.GA)
```

```{r}
ggplot(data=rang_imps_ordered,aes(x=rowname, y= rang_imps))+geom_point()+xlim(0,1000)



```

```{r}
data_train_GA_rf_top100[1:5,1:10]
```



Okay. Still not so hot. I think we've got a crappy feature set, so how about better feature engineering? 

Shall we try RFE? 

1a) LASSO, bitches
3. SVM for classification and regression
4. LDA for classification
5. Random Forest for GA and TERM

6. RFE on larger dataset


Lets try lasso: 


```{r}
lasso_control <-  trainControl(method = "repeatedcv", repeats = 10, number = 5)
```


```{r}
lasso_rf_100 <- caret::train(GA~., data=data_train_GA_rf_top100, method= "lasso", trControl = lasso_control)
lasso_rf_1000 <- caret::train(GA~., data=data_train_GA_rf_top1000, method= "lasso", 
trControl = lasso_control)
```



```{r}
lasso_GA_100 <- caret::train(GA~., data=data_train_GA_lc_top100, method= "lasso", trControl = lasso_control)
lasso_GA_1000 <- caret::train(GA~., data=data_train_GA_lc_top1000, method= "lasso", trControl = lasso_control)
```


```{r}
lasso_pca_100  <- caret::train(GA~., data=data_train_GA_pca_100, method= "lasso", trControl = lasso_control)

```

```{r}

lasso_GA_100
lasso_GA_1000
lasso_pca_100
```


Lasso works pretty well it seems, so how about elastic net? 


okay elastic net and lasso and ridge properly 

```{r}
data_train_GA_rf_label <- data_train_GA_rf_top100 %>% select("GA")
#data_train_rownames_term_label <-data_train_rownames_term %>% select("term")
data_train_GA_rf_100_no_label <- data_train_GA_rf_top100 %>% select(-"GA")
data_train_GA_rf_500_no_label <- data_train_GA_rf_top500 %>% select(-"GA")
data_train_GA_rf_1000_no_label <- data_train_GA_rf_top1000 %>% select(-"GA")
GA_rf100_no_lab_mat <- data.matrix(data_train_GA_rf_100_no_label, rownames.force = TRUE)
GA_rf500_no_lab_mat <- data.matrix(data_train_GA_rf_500_no_label, rownames.force = TRUE) 
GA_rf1000_no_lab_mat <- data.matrix(data_train_GA_rf_1000_no_label, rownames.force = TRUE)

GA_rf_lab_mat <- data.matrix(data_train_GA_rf_label, rownames.force = TRUE)

```

```{r}
ridge_fit_rf100 = cv.glmnet(x = scale(GA_rf100_no_lab_mat), y = GA_lab_mat, alpha = 0)
ridge_fit_rf500 = cv.glmnet(x = scale(GA_rf500_no_lab_mat), y = GA_lab_mat, alpha = 0)
ridge_fit_rf1000 = cv.glmnet(x = scale(GA_rf1000_no_lab_mat), y = GA_lab_mat, alpha = 0)
```


```{r}
ridge_fit_rf100 
ridge_fit_rf500 
ridge_fit_rf1000
```


```{r}
lasso_fit_rf100 = cv.glmnet(x = scale(GA_rf100_no_lab_mat), y = GA_lab_mat, alpha = 1)
lasso_fit_rf500 = cv.glmnet(x = scale(GA_rf500_no_lab_mat), y = GA_lab_mat, alpha = 1)
lasso_fit_rf1000 = cv.glmnet(x = scale(GA_rf1000_no_lab_mat), y = GA_lab_mat, alpha = 1)
```


```{r}
lasso_fit_rf100
lasso_fit_rf500
lasso_fit_rf1000
```

```{r}
sqrt(33.13)
```

```{r}
elastic_control <-  `train`Control(method = "cv", number = 5)
```



```{r}
elastic_rf_100 <- caret::train(GA~., data=data_train_rf_lc_top100, method= "glmnet", trControl = elastic_control, tuneLength=10)
elastic_rf_pca_100 <- caret::train(GA~., data=data_train_rf_pca_100, method= "glmnet", trControl = elastic_control, tuneLength=10)
```

```{r}
elastic_GA_100
elastic_GA_pca_100
```



# Here we try LDA


```{r}
san_LDA <- train(term~., data=data_san_term, method= "lda", trControl = control_rf_term)
```

# SVM

# Linear

#radial


#ranger
```{r}
rf_san_ranger_term <-train(term~., data= data_san_term, method = "ranger", trControl = control_rf_term)
```


```{r}
rf_san_ranger_GA <-train(GA~., data= data_san_GA, method = "ranger", trControl = control_rf_GAw)

```













#```{r} 
#write.table(data_df, "GAtext.txt", sep = "\t", row.names = FALSE, col.names = TRUE, quote=F)
#```

#Feature selection 

#Todo: feed PCA'd data to lasso
#Todo: Consider ICA
#Todo: Run random forest and SVM, consider LDA

#randomforest
# still w/ stack overflow on 16 cores at 4 per. ..



#feature selection with univariate filtering

#fselector 
#using correlation

```{r}
lctest <- linear.correlation(GA~., data_san_GA)
```

```{r}
str(lctest)
```



#lets try boruta

```{r}
boruta_test <- Boruta(as.factor(term)~., data_san_term, doTrace=2 )
```

random forest: 


```{r}
rf_san_GA <- train(GA~., data= data_san_GA, method = "rf", trControl = control_rf_GA)

```

```{r}
rf_imp<- varImp(rf_san_term, scale= FALSE)
```

```{r}
rfitest2 <- rownames_to_column(rf_imp)
rfitest2 <- rename(rfitest2, gene = rowname)
rfitest_ordered <- rfitest2 %>% arrange(desc(attr_importance))
rfi_top100 <- rfitest_ordered[1:100,]
rfi_top200 <- rfitest_ordered[1:200,]
rfi_top500 <- rfitest_ordered[1:500,]
#rfi_top1000 <- rfitest_ordered[1:1000,]
```


```{r}
rfi_100 <- rfi_top100$gene
rfi_200 <- rfi_top200$gene
rfi_500 <- rfi_top500$gene
#rfi_1000 <- rfi_top1000$gene

data_san_term_rfi_top100 <- subset(data_san_term, select= rfi_100)
data_san_term_rfi_top200 <- subset(data_san_term, select= rfi_200)
data_san_term_rfi_top500 <- subset(data_san_term, select= rfi_500)
#data_san_term_rfi_top1000 <- subset(data_san_term, select= rfi_1000)


```

#failed

```{r}
rf_imp <- random.forest.importance(term~., data= data_san_term, importance.type = 1)
```


```{r}
class(data_san_term)
```



RFE? or after? 




### trying feser
This seems to be very good in that it dumps a huge number of things. We're down 3k features with this alone. 

The next question is whether we should remove minimal variance 

```{r}
output <- filter.corr(features= GA_no_lab_mat, class = GA_lab_mat, mincorr = 0.1)
```







```{r}
control_rf_term <- trainControl(method="repeatedcv", number = 5, repeats = 5, classProbs = T, savePredictions = T)
control_rf_GA <- trainControl(method="repeatedcv", number = 5, repeats = 5)

```


#regression 
#Logistic

OKAY! BUILD MODELS!

1. Regression against age
```{r}

```
1a) LASSO, bitches

2. Logistic Regression for term 

3. SVM 
4. LDA
5. Random Forest for GA and TERM
6. RFE on larger dataset

```{r}
san_glm_log <- train(term~., data=data_san_term, method= "glm", family= "binomial", trControl = control_rf_term)

```

```{r}
san_glm_lin <- train(GA~., data=data_san_GA, method= "glm", family= "gaussian", trControl = control_rf_term)

```


#random forest

```{r}
rf_imp<- varImp(rf_san_term, scale= FALSE)
```



```{r}
rfitest2 <- rownames_to_column(rf_imp)
rfitest2 <- rename(rfitest2, gene = rowname)
rfitest_ordered <- rfitest2 %>% arrange(desc(attr_importance))
rfi_top100 <- rfitest_ordered[1:100,]
rfi_top200 <- rfitest_ordered[1:200,]
rfi_top500 <- rfitest_ordered[1:500,]
#rfi_top1000 <- rfitest_ordered[1:1000,]
```


```{r}
rfi_100 <- rfi_top100$gene
rfi_200 <- rfi_top200$gene
rfi_500 <- rfi_top500$gene
#rfi_1000 <- rfi_top1000$gene

data_san_term_rfi_top100 <- subset(data_san_term, select= rfi_100)
data_san_term_rfi_top200 <- subset(data_san_term, select= rfi_200)
data_san_term_rfi_top500 <- subset(data_san_term, select= rfi_500)
#data_san_term_rfi_top1000 <- subset(data_san_term, select= rfi_1000)


```

using caret's built in sbf, not great
```{r}
sbf_filter_control <- sbfControl(functions = rfSBF, method ="repeatedcv", number = 5, repeats= 5)
testsbf <- sbf(term~.,data = data_san_term, sbfControl = sbf_filter_control)
```


```{r}

```


#LDA 
# not normalized or scaled
#remove collinearity?





```

```{r}
rfterm = ranger(GA~., data = data_train_GA[1:3,],  mtry= 1000, num.trees= 1000, min.node.size = 10000, save.memory = TRUE)
save(rftrainGA, file= "rftrain.RData")

```


#LDA on term vs. 


```{r}
data_train_term_rn <- data_train_term %>% column_to_rownames(.,var = "SampleID")
```


```{r}
lda_test <- lda(term ~., data= data_train_term_rn)
```



4. Test model on test set 
```{r}
#predict the species of the test data
LDA_predict <- predict(LDA, newdata=test_classifier)
confusionMatrix(LDA_predict$class, reference = test_classifier$Species)
```

#svm

```{r}

```



New todo list: 

-feature selection
- use built in caret univariate filtration
- pca
- other? 

- tuning grid for ranger

Clip low variance features .
Consider Geseq

Demoted code chunks

Attaches lc correlation scores to GA data frame


```{r}
lctest_transpose <- as.data.frame(t(lctest))
dummyGA <- data.frame("GA" =0)
lctest_transpose <- cbind(dummyGA, lctest_transpose)
data_san_GA_lc <- rbind(lctest_transpose, data_san_GA)
data_san_GA_lc[1:20,1:20]
```

using caret's built in sbf, not great
```{r}
filter_control <- sbfControl(functions = lmSBF, method ="repeatedcv", number = 5, repeats= 5)
testsbf <- sbf(GA~.,data = data_san_GA, sbfControl = filter_control)
```

```{r}
testsbf
```



```{r}
GA_no_lab_mat <- data.matrix(data_san_GA_no_label, rownames.force = TRUE)
GA_lab_mat <- data.matrix(data_san_GA_label, rownames.force = TRUE)
```

```{r}
rownames(data_san_GA_no_label)
```


```{r}
rownames(GA_no_lab_mat)
```


```{r}
nzv <- nearZeroVar(GA_no_lab_mat, saveMetrics = TRUE)
```



```{r}
str(colvar)
```

```{r}
colvar <-  as.data.frame(colVars(GA_no_lab_mat))
colvar <-  rownames_to_column(colvar)
names(colvar)[names(colvar) == "colVars(GA_no_lab_mat)"] <- "variance"
colvar <- colvar[order(-colvar$variance),]
```




```{r}
data_san_GA_label <- data_san_GA %>% select("GA")
data_san_term_label <-data_san_term %>% select("term")
```

```{r}
data_san_GA_no_label <- data_san_GA %>% select(-"GA")
data_san_term_no_label <- data_san_term %>% select(-"term")
```



#sanity check

```{r}
data_san_term <- data_train_term[,1:1000]
data_san_GA <- data_train_GA[,1:1000]
```

```{r}
data_san_term <- data_san_term %>% select(-"SampleID")
data_san_GA <- data_san_GA %>% select(-"SampleID")

```

```{r}
#convert numeric scoring to factor for term
data_san_term$term <- factor(data_san_term$term)
data_san_term <- data_san_term %>% mutate(term = ifelse(term==1, "yes", "no"))
```

```{r}
data_san_GA_label <- data_san_GA %>% select("GA")
data_san_term_label <-data_san_term %>% select("term")
```

```{r}
data_san_GA_no_label <- data_san_GA %>% select(-"GA")
data_san_term_no_label <- data_san_term %>% select(-"term")
```
